
\chapter{Multidimensional Visual Analysis (MVA)}

\label{chap:MVA}

Multidimensional Visual Analysis (MVA) approaches are methods and
techniques used to analyze and understand complex data sets using
visual representations. These approaches typically involve the use of
specialized software or tools that allow analysts to create and
manipulate graphical representations of the data in order to uncover
patterns, trends, and relationships. In this section, we will review
some popular MVA approaches.



\section{Scatter Plots}

A scatter plot is a type of graph that is used to display the
relationship between two numerical variables, to identify any
potential trends or patterns in the data, and to identify outliers. It
uses dots or markers to represent the values of the two variables, and
position of each dot on the graph indicates the value of the two
variables in a single observation.

To create a scatter plot, the values of one variable are plotted on
the x-axis (horizontal axis) and the values of the other variable are
plotted on the y-axis (vertical axis). The resulting graph will show a
set of dots, with each dot representing a single observation. If there
is a positive relationship between the two variables, the dots will
tend to form a diagonal line that slopes upwards from left to
right. If there is a negative relationship, the dots will tend to form
a diagonal line that slopes downwards from left to right. If there is
no relationship between the two variables, the dots will be scattered
randomly across the graph.





\section{Similarity Maps}

Similarity maps are projections of high-dimensional datasets to two
(or sometimes three) dimensions. Projection techniques are used to
reduce the number of dimensions in the data, while attempting to
preserve distances between items as far as possible. Items which are
close in the high-dimensional space should also be close in the
resulting two-dimensional projection space. Such projections can be
further split into \emph{linear} and \emph{non-linear} projections.


\subsection{Principal Component Analysis (PCA)}

Principal Component Analysis (PCA) is a linear projection. PCA uses
linear algebra to identify the underlying dimensions or factors in the
data and project the data onto a lower-dimensional space
\parencite{abdi2010principal}.


\subsection{Multi-Dimensional Scaling (MDS)}

Multi-Dimensional Scaling (MDS) is a non-linear projection. MDS uses a
distance metric to preserve the distances between data points in the
high-dimensional space and project the data onto a lower-dimensional
space \parencite{morrison2003fast}.


\subsection{t-Distributed Stochastic Neighbor Embedding (t-SNE)}

t-distributed stochastic neighbor embedding (t-SNE) is a non-linear
projection. t-SNE uses a probabilistic model to preserve the local
structure of the data in the high-dimensional space and project the
data onto a lower-dimensional space \parencite{van2008visualizing}.


% UMAP





\section{Parallel Coordinates}

Parallel coordinates are a type of chart that is used to visualize
multi-dimensional data. In this type of chart, each data point is
represented by a vertical line that extends across multiple axes. The axes
are typically arranged in parallel. Each axis represents a different
variable, and the position of a point on that axis indicates the value of
that variable for the data point. This allows multiple variables to be
compared and analyzed simultaneously \parencite{inselberg1990parallel}.







\section{Brushing and Linking}

Brushing and linking are techniques used in multidimensional visual
analysis to allow the user to interact with a visualization and
explore data in greater depth.

Brushing refers to the process of selecting data points or regions in
one visualization and highlighting those data points in other
visualizations. This allows the user to see how the data points or
regions of interest are related to other variables in the dataset.

Linking refers to the process of synchronizing the views of multiple
visualizations, such that a change made to one visualization is
reflected in the other visualizations. This allows the user to explore
the data from different perspectives and understand how different
variables are related to one another.

Both brushing and linking are useful for helping users to identify
patterns and relationships in the data and for facilitating the
process of data exploration and analysis.






\section{Grouping and Labelling}

Grouping and labeling data is a fundamental step in the ML process.
Grouping refers to the process of identifying and separating similar data
points, while labeling refers to the process of providing relevant
information about the data. Grouping and labelling data helps to organize
and structure the data, making it easier to understand and work with. It
allows for more accurate and meaningful analysis of the data, as similar
data points can be grouped together and analyzed in relation to one
another. It also allows the ML model to understand the context of the data
and make more accurate predictions, and enables the interpretability of
the model, by providing meaningful information about the data and how it
is being used.

\subsection{Manual Grouping}

Manual grouping and labeling refers to the process of organizing and
providing relevant information about data manually, typically done by a
human. The process of manual grouping and labeling can be time-consuming
and labor-intensive.




\subsection{Automated Clustering}

Cluster analysis is a type of statistical technique used to identify
groups of similar objects within a data set. It is a commonly used
method for exploratory data analysis, and is often used as a way to
gain insight into the underlying structure of the data. In cluster
analysis, the data is divided into groups, or clusters, based on the
similarity of the objects within each group. This allows analysts to
identify common patterns and trends within the data, and to gain a
better understanding of the relationships between the objects in the
data set \parencite{duran2013cluster}.



