%----------------------------------------------------------------
%
%  File    :  survey-intro.tex
%
%  Author  :  Keith Andrews, IICM, TU Graz, Austria
% 
%  Created :  27 May 1993
% 
%  Changed :  16 Nov 2010
% 
%----------------------------------------------------------------


\chapter{Introduction}

\label{chap:Intro}



Multidimensional Visual Analysis (MVA) is a field that focuses on the
use of visual representations to explore and analyze complex data
sets. This approach aims to provide a more intuitive and interactive
way to understand and extract insights from large and complex
datasets.

One of the main challenges in MVA is the effective representation of
high-dimensional data in a way that is meaningful and intuitive for
the user. To address this challenge, a wide range of visual encodings
and interaction techniques have been developed, including scatter
plots, parallel coordinates, and multidimensional scaling.

One of the key benefits of MVA is its ability to reveal patterns and
relationships in the data that may not be apparent through traditional
statistical analysis methods. This is particularly useful in the
exploratory phase of data analysis, where the aim is to gain a better
understanding of the data and identify potential areas of interest for
further investigation.

In addition to its use in exploratory data analysis, MVA also has a
range of applications in areas such as data mining, machine learning,
and business intelligence. By enabling users to interact with and
visualize data in a more intuitive way, it can help to improve
decision-making and facilitate the discovery of insights and trends
that may not have been identified through other means.

Overall, MVA is an important tool for understanding and making sense
of complex data, and has the potential to greatly enhance our ability
to extract knowledge and insights from large and complex datasets.

This survey reviews multiple popular MVA approaches and popular MVA
software. See Chapter~\ref{chap:MVA} for a review of MVA approaches and
Chapter~\ref{chap:MVATools} for a review and comparison of MVA tools.

